---
title: "Spatial Analysis of Oil Spills In California"
description: |
  Mapping spatial data using R.
author:
  - name: Thomas Wheeler
date: 02-25-2021
output:
  distill::distill_article:
    self_contained: false
    theme: space
    code_folding: hide
categories:
  - interactive-data-visualization
  - spatial-analysis
---


### Overview

This project conducts an exploratory analysis of oil spill incidents in California. The first tab is an interactive graph of oil spill locations throughout California and the second visualizes cumulative counties of oil spills by county to better understand which counties are most impacted by oil spills.

### Data

The Office of Spill Prevention and Response (OSPR) Incident Tracking Database is a statewide oil spill tracking information system. The data are collected by OSPR Field Response Team members for Marine oil spills and by OSPR Inland Pollution Coordinators and Wardens for Inland incidents.

Data source: **Office of Spill Prevention and Response (OSPR) Incident Tracking Database** [CA.gov](https://gis.data.ca.gov/datasets/7464e3d6f4924b50ad06e5a553d71086_0/data)

### Packages Used in Analysis

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap)
library(gstat)
library(stars)
```

## {.tabset}

### Oil Spill Locations in California

This interactive map provides an overview of all oil spill locations recorded by the Office of Spill Prevention and Response.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#read in California counties layer
ca_counties <- read_sf(here("_posts", "2021-02-08-welcome", "welcome_files", "data"), layer = "CA_Counties_TIGER2016") %>%
  clean_names()

#read in California counties layer
oil_spills <- read_sf(here("_posts", "2021-02-08-welcome", "welcome_files", "data"), layer = "Oil_Spill_Incident_Tracking_%5Bds394%5D") %>%
  clean_names()
```

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#make tmap interactive
tmap_mode("view")

#plot interactive map
tm_shape(oil_spills) +
  tm_dots()
```

### Count of Oil Spills by County

This static map aggregates all oil spill incidents recorded by county.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#join oil_spills with california counties layer
ca_oil_spills <- ca_counties %>% 
  st_join(oil_spills)

#count of values in ca counties
ca_counts <- ca_oil_spills %>% 
  count(name)

#create finalized static cloropleth of count of oil spills by county
ggplot(data = ca_counts) +
  geom_sf(aes(fill = n), color = "white", size = 0.1) +
  scale_fill_gradientn(colors = c("lightgray","orange","red")) +
  theme_minimal() +
  labs(fill = "Number of oil spills")
```
